### 2016-07-12 cross_compilation.tex
Working on kodo cross-compiling

\subsection{Cross compile example}

The following example will provide a very basic example of 1) how to cross compile
a \texttt{hello\_world} c++ example to the \ac{Raspi}'s ARM architecture
and 2) how to copy and execute the binary in a \ac{Raspi} using \ac{SCP} and
\ac{SSH}.

Create the file \texttt{hello\_world.cpp} with the following content using your
favorite text editor:

\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\${CODEDIR}/hello\_world.cpp}@>
<@\textcolor{gray}{
---------------------------------------------------------------}
\Reactivatenumber @>
#include <iostream>

int main()
{
    std::cout << "Hello World!" << std::endl;
    return 0;
}
\end{lstlisting}
\FloatBarrier

Save \texttt{hello\_world.cpp} and compile it for \ac{Raspi}:

\begin{lstlisting}[]
$ raspberry-gxx49-arm-g++ hello_world.cpp -o hello_world
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

This should produce a binary \texttt{hello\_world}. Copy it to a \ac{Raspi}
and test it. We will use \ac{SCP} to copy the binary file to one of the \ac{Raspi}s.

The default username for our Raspbian lite distribution is "pi" and password is "raspberry"

\begin{lstlisting}[]
$ scp main pi:<RASP_IP>:~/
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

If you do not know the \ac{IP} address of a \ac{Raspi} in your network, you can
connect a monitor to it and run the following command after you have logged in:
\begin{lstlisting}[]
pi@rasp01:~ $ ifconfig
eth0      Link encap:Ethernet  HWaddr b8:27:eb:72:77:54
          inet addr:192.168.87.107  Bcast:192.168.87.255  Mask:255.255.255.0
          inet6 addr: fe80::e0a5:38f3:6f82:bc79/64 Scope:Link
          UP BROADCAST RUNNING MULTICAST  MTU:1500  Metric:1
          RX packets:1537 errors:0 dropped:0 overruns:0 frame:0
          TX packets:445 errors:0 dropped:0 overruns:0 carrier:0
          collisions:0 txqueuelen:1000
          RX bytes:259117 (253.0 KiB)  TX bytes:52551 (51.3 KiB)

\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

After the executable has been copied to a \ac{Raspi}. Then, \ac{SSH} to it:

@@

We can list the directory content after we have logged into the \ac{Raspi} and
see that \texttt{hello\_world} is there:

\begin{lstlisting}[]
pi@rasp07:~ $ ls
hello_world  rasp_config
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Now, simply execute the \texttt{hello\_world} to confirm that cross compiling
\texttt{hello\_world} worked properly:

\begin{lstlisting}[]
pi@rasp07:~ $ ./hello_world
Hello World!
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\subsection{Cross compile Kodo}

Kodo is a C++11 network coding library. It is available through bindings in
a variety of programming languages.
%There exists a number of bindings for
%Kodo to other popular programming languages.
This procedure will present how to configure the Kodo C++ bindings (Kodo-cpp)
to cross compile applications that can run in \ac{Raspi}. More details are
provided in the code documentation~\cite{kodocppdoc}.

%This procedure will go through
%how to setup Kodo-cpp to cross compile applications that can run in \ac{Raspi}. Kodo-cpp provides a
%simple interface to the underlying C++11 code that exists in the libraries
%Kodo-core and Kodo-rlnc.

To use Kodo, it is required to obtain a research- and
education-friendly license.
It can be be obtained from~\cite{steinwurflicenselink} and is free
for research and educational purposes.

%A more detailed procedure of setting up Kodo-cpp than provided below is available at

\begin{enumerate}

\item Start by creating a directory to hold the code:
\begin{lstlisting}[]
$ CODEDIR="${HOME}/code"
$ mkdir -p ${CODEDIR}
$ cd ${CODEDIR}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\item Clone the Kodo repository and change directory into the repository:
\begin{lstlisting}[]
$ git clone https://github.com/steinwurf/kodo-cpp.git
$ cd kodo-cpp
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\item Configure Kodo to build executables for the \ac{ARM} architecture using the \ac{Raspi} toolchain:
\begin{lstlisting}[]
$ python waf configure --cxx_mkspec=cxx_raspberry_gxx49_arm
...
'configure' finished successfully (0.620s)
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\item Build executables:
\begin{lstlisting}[]
$ python waf build
...
'build' finished successfully (2m22.918s)
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\item Make shared library:
\begin{lstlisting}[]
$ python waf install --install_shared_libs --install_path="./shared_test"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\item Copy shared library to a \ac{Raspi}'s home directory (Alternatively, Kodo can also generate static libraries):
\begin{lstlisting}[]
$ scp -r shared_test/include shared_test/libkodoc.so pi@<RASP_IP>:~/
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\item Copy a binary using Kodo shared library to the \ac{Raspi}'s home directory:
\begin{lstlisting}[]
$ scp -r shared_test/encode_decode_simple pi@<RASP_IP>:~/
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\item Execute binary:
\begin{lstlisting}[]
$ ssh pi@<RASP_IP> ./encode_decode_simple
Data decoded correctly
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

\end{enumerate}

>>>>

\subsection{Cross-compile Example}
The following shows: (i) How to cross compile the classic \texttt{hello\_world}
C++ example for the \ac{Raspi} ARM architecture and (ii) how to copy and
execute the binary in a \ac{Raspi} using \ac{SCP} and \ac{SSH}. First, we
create the file \texttt{hello\_world.cpp}. For simplicity, we create it
in the directory where we stored the \texttt{fabfile.py} file
with the following content using any text editor:

\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\${CODEDIR}/hello\_world.cpp}@>
<@\textcolor{gray}{
-----------------------------------------------}
\Reactivatenumber @>
#include <iostream>

int main()
{
    std::cout << "Hello World!" << std::endl;
    return 0;
}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

We save the previous file and compile it for \ac{Raspi} in the testbed
administrator \ac{PC} by doing:

\begin{lstlisting}[]
$ ${TOOLCHAINBINARY} hello_world.cpp -o hello_world
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

This should produce a binary \texttt{hello\_world} that should only be
useful in the \ac{Raspi}. We copy it to a \ac{Raspi} to test it. We will
use \ac{SCP} to copy the binary file to one of the \ac{Raspi}s. Still, we
for example might use Fabric if we are interested in deploying a compiled
binary for many \ac{Raspi}s.

\begin{lstlisting}[]
$ scp hello_world pi@<RASP_IP>:~/
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

% If you do not know the \ac{IP} address of a \ac{Raspi} in your network, you can
% connect a monitor to it and run the following command after you have logged in:
% \begin{lstlisting}[]
% pi@rasp01:~ $ ifconfig
% eth0      Link encap:Ethernet  HWaddr b8:27:eb:72:77:54
%           inet addr:192.168.87.107  Bcast:192.168.87.255  Mask:255.255.255.0
%           inet6 addr: fe80::e0a5:38f3:6f82:bc79/64 Scope:Link
%           UP BROADCAST RUNNING MULTICAST  MTU:1500  Metric:1
%           RX packets:1537 errors:0 dropped:0 overruns:0 frame:0
%           TX packets:445 errors:0 dropped:0 overruns:0 carrier:0
%           collisions:0 txqueuelen:1000
%           RX bytes:259117 (253.0 KiB)  TX bytes:52551 (51.3 KiB)

% \end{lstlisting}
% \FloatBarrier
% \vspace{-5mm}

After the executable has been copied to the \ac{Raspi}. We login through
\ac{SSH} to it:

@@

We can list the directory content after we have logged into the \ac{Raspi} and
verify that the compiled \texttt{hello\_world} binary is there:

\begin{lstlisting}[]
pi@<RASP_IP>:~ $ ls
hello_world  rasp_config
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Finally, we simply execute the \texttt{hello\_world} to confirm that
the cross-compiling of \texttt{hello\_world} worked properly:

\begin{lstlisting}[]
pi@<RASP_IP>:~ $ ./hello_world
Hello World!
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\subsection{Cross-compile Kodo}

As we originally mentioned, Kodo is a C++11 network coding library that
permits to implement network coding functionalities by allowing any network
protocol designer to use and test the primitive encoding, decoding and
recoding operations of \ac{RLNC}. In this way, a designer only needs to
focus on the design and test of a network coding-based protocol. Kodo is
available through programming bindings for a variety of popular
programming languages. This procedure will present how to configure the
Kodo C++ bindings \texttt{kodo-cpp} to cross-compile applications that can
run in \ac{Raspi}. \texttt{kodo-cpp} provides a simple interface to the
underlying C++11 code that exists in the libraries \texttt{kodo-core} for
the object structure and \texttt{kodo-rlnc} for the \ac{RLNC} codec
implementation. More details about Kodo are provided in the code
documentation~\cite{kodocppdoc}.

To use Kodo for research, it is required to obtain a research free license.
To do this, only a request form needs to be filled
in~\cite{steinwurflicenselink} and wait for grant access. Once
the access for Kodo has been granted, the source code can be pulled from
its Git repositories to be compiled. Thus, assuming that testbed
administrator already has access, we clone the Kodo repository locally
in \texttt{\${CODEDIR}} and change directory into the repository by doing:

\begin{lstlisting}[]
$ cd ${CODEDIR}
$ git clone git@github.com:steinwurf/kodo-cpp.git
$ cd kodo-cpp
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Configure Kodo to build executables for the \ac{ARM} architecture using
the \ac{Raspi} toolchain:
\begin{lstlisting}[]
$ python waf configure --cxx_mkspec=cxx_raspberry_gxx49_arm
...
'configure' finished successfully (X.XXXs)
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Build executables:
\begin{lstlisting}[]
$ python waf build --run_tests
...
'build' finished successfully (2m22.918s)
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Make shared library:
\begin{lstlisting}[]
$ python waf install --install_shared_libs --install_path="./shared_test"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Copy shared library to a \ac{Raspi}'s home directory (Alternatively, Kodo can also generate static libraries):
\begin{lstlisting}[]
$ scp -r shared_test/include shared_test/libkodoc.so pi@<RASP_IP>:~/
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Copy a binary using Kodo shared library to the \ac{Raspi}'s home directory:
\begin{lstlisting}[]
$ scp -r shared_test/encode_decode_simple pi@<RASP_IP>:~/
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Execute binary:
\begin{lstlisting}[]
$ ssh pi@<RASP_IP> ./encode_decode_simple
Data decoded correctly
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@
